---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Peer assignment 2 
===================

This assignment is about analyzing the data from a fitbit.

### Load data 
```{r, echo = TRUE, results ='hide'}
df <- read.csv("activity.csv")
```

### Pre process data
```{r Total steps each day, echo = TRUE, results='hide'}
total_steps <- with(df, aggregate(steps, by = list(date), sum, na.rm =TRUE))
names(total_steps) <- c("date", "total.steps")
```
### What is mean total number of steps taken per day?
### 1. Make a histogram of the total number of steps taken each day
```{r histogram}
hist(total_steps$total.steps, col = "green", xlab = "Total steps taken each day", ylab = "Frequency", main = "Total steps taken each day")
```

### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
mean<- mean(total_steps$total.steps)
median <- median(total_steps$total.steps)
```

### Mean total number of steps each day is `r mean` and median is `r median`  

### What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r Mean steps}
library(ggplot2)
mean_steps <- with(df, aggregate(steps, by = list(interval), mean, na.rm = TRUE))
names(mean_steps)<- c("interval", "steps")
qplot(interval, steps, geom = "line", data = mean_steps, xlab = "Time interval", ylab = "Avg number of steps across all days")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r max_interval}
max_interval_df <- subset(mean_steps, mean_steps$steps == max(mean_steps$steps))
maxInterval<- max_interval_df[[1]]
```
Max number of steps across all days is during `r maxInterval`th interval.

### Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
num_na <- sum(is.na(df))
```
Number of NA rows is `r num_na`

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Strategy is to fill the missing values with the median steps for the respective interval across all days. 

```{r strategy for missing values}
msteps <-  with(df, aggregate(steps, by = list(interval), median, na.rm = TRUE))
names(msteps) <- c("interval", "steps")

mergeddf<- merge(df, msteps, by = "interval")
mergeddf<- mergeddf[order(mergeddf$date),]
```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r new data set, results='hide'}
for ( i in 1:nrow(mergeddf))
{
    if(is.na(mergeddf[i, ]$steps.x))
      mergeddf[i,]$steps.x = mergeddf[i,]$steps.y
}
```

4. Make a histogram of the total number of steps taken each day and  Do these values differ from the estimates from the first part of the assignment? 

```{r}
total_steps<- aggregate(mergeddf$steps.x, by = list(mergeddf$date), sum)
names(total_steps)<- c("interval", "steps")

hist(total_steps$steps, col = "green", xlab = "Total steps taken each day", ylab = "Frequency", main = "Total steps taken each day with median missing values")
```

Calculate and report the mean and median total number of steps taken per day.
```{r}
mean<- mean(total_steps$steps)
median <- median(total_steps$steps)
```
### . Mean total number of steps each day is `r mean` and median is `r median`.

What is the impact of imputing missing data on the estimates of the total daily number of steps?  
- Mean value differes slightly but median is the same since we used the median value for imputing missing values.  



### Are there differences in activity patterns between weekdays and weekends?

For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
mergeddf$day <- weekdays(as.Date(mergeddf$date))
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
par(mfrow=c(2,1))

week_days <- subset(mergeddf, mergeddf$day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"))
weekday_steps <- with(week_days, aggregate(steps.x, by = list(interval), mean))

weekend_days <- subset(mergeddf, mergeddf$day %in% c("Saturday","Sunday"))
weekend_steps <- with(weekend_days, aggregate(steps.x, by = list(interval), mean))

names(weekday_steps)<- c("interval", "avg_steps")
names(weekend_steps)<- c("interval", "avg_steps")


qplot(interval, avg_steps , geom = "line", data = weekday_steps, xlab = "Time interval", ylab = "Avg number of steps across week days")
qplot(interval, avg_steps , geom = "line", data = weekend_steps, xlab = "Time interval", ylab = "Avg number of steps across weekend days")
```